AN EFFICIENT LOCALIZED MODEL ORDER REDUCTION FRAMEWORK FOR THE SHAPE OPTIMIZATION OF ADDITIVELY MANUFACTURED LATTICE STRUCTURES\break ECCOMAS CONGRESS 2024

### 2024-08-19 paper/paper.tex
[sec 2.2] initial commit

The contributions of the present work are given as follows.
First, as references~\cite{Smetana2016Optimal,Taddei2018Localization} do not make use of range approximation via random sampling, suitable training strategies to construct local approximation spaces for parameterized linear problems via random sampling are investigated.
In particular, the approach given in~\cite{Taddei2018Localization}, identified as a \textit{distributed approximate POD} (see also~\cite{Himpe2018Hierarchical}), is compared to a \textit{heuristic randomized range finder} developed in this work.
Second, a framework for the shape optimization of lattice structures is proposed.
It combines the aforementioned algorithms to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
Furthermore, the global approximation space is constructed from the local spaces using the generalized finite element method (GFEM)~\cite{Babuska2004Generalized}.

The rest of the article is organized as follows.
In~\cref{sec:method}, the building blocks of the shape optimization framework are described.
In particular, the aforementioned two different strategies for the construction of local approximation spaces are presented.
\Cref{sec:numerical experiments} comprises the numerical experiments.
Based on the example of a graded concrete slab, the methods for the construction of local approximation spaces described in~\cref{sec:Randomized range finder and POD,sec:Heuristic randomized range finder} are compared, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
Finally, conclusions are given in~\cref{sec:conclusions}.

\section{METHOD}%
\label{sec:method}
The method proposed in this article is a localized pMOR framework for the shape optimization of lattice structures.
First, the auxiliary problem to model geometrical parameterizations is introduced in~\cref{sub:Auxiliary Problem}.
Second, the construction of local approximation spaces is described in~\cref{sub:Construction of local approximation spaces}.
Herein, the parametric oversampling problem and two different strategies to approximate the range of the corresponding parametric transfer operator are discussed.
Third, it is outlined how a global approximation is obtained from local approximation spaces via the GFEM.
Finally, the necessity and use of hyperreduction (in the form of empirical interpolation) to ensure an online-efficient ROM is detailed in~\cref{sub:Online efficiency}.

\subsection{Auxiliary Problem} % (fold)
\label{sub:Auxiliary Problem}
The approach adopted in this work to facilitate geometrical parametrizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\bm\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\bm\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\bm\mu}$.
In the context of shape optimization of lattice structures, our objective is to determine the mapping $\Phi_{\bm\mu}$ for a single unit cell, see~\cref{fig:transformationmap}, and use it to describe the change in geometry of each unit cell throughout the structure.
An auxiliary problem based on the equations of linear elasticity is solved to obtain such domain transformations, following the approach outlined in~\cite{Guo2022Learning}, and briefly repeated here for completeness.

@@

Given the transformation displacement $\bm{d}(\bm{x}_{\mathrm{p}};\bm\mu)$, the variational formulation is formulated over the parameter-independent parent domain $\varOmega^{\mathrm{p}}$ instead of the parameter-dependent physical domain $\varOmega^{\bm\mu}$.
This introduces the parameter dependence in the variational formulation as shown in the following subsections, but has the advantage that all computations (integrations) can be carried out over the fixed parent domain and thus no re-meshing is required.
However, for each new parameter value $\bm\mu$ for which the model is to be evaluated, we first need to solve the auxiliary problem.
Nevertheless, this does not pose a problem, since the solution of the auxiliary problem is well amenable to acceleration via established MOR techniques and it is referred to~\cite{Guo2022Learning} for further details.

@@

\subsection{Construction of local approximation spaces} % (fold)
\label{sub:Construction of local approximation spaces}

% - setting, global domain = union of non overlapping subdomains = unit cells
% - introduction of coarse grid where each cell corresponds to unit cell subdomain
% - goal of constructing local spaces for only parts of the domain
% - definition of oversampling problem and transfer operator
% - the oversampling problem should be formulated for union of parent unit cells
% - weak formulation is not required here? Should be first introduced in section 2.3?
% - comment on number of archetypes/configurations

@@

\subsubsection{Heuristic randomized range finder} % (fold)
\label{sec:Heuristic randomized range finder}

\begin{algorithm}
    \caption{HRRF caption.}%
    \label{algo:hrrf}
    \begin{algorithmic}[1]
        \Function{HeuristicRangeApproximation}{$\bm{T}$, $\texttt{tol}$, $n_t$, $\varepsilon_{\mathrm{algofail}}$, $N_{\mathrm{train}}$, $N_{\mathrm{test}}$}
    \Statex{\textbf{Input:} Operator $\bm{T}_{\bm\mu}$, target tolerance $\texttt{tol}$, number of testvectors $n_t$, maximum failure probability $\varepsilon_{\mathrm{algofail}}$, number of parameter samples $N_{\mathrm{train}}$, number of parameter samples $N_{\mathrm{test}}$}
    \Statex{\textbf{Output:} space $X^n$}

    \State{$\bm{B}$ $\gets\;\emptyset$}\Comment{initialize basis}
    \State{$\bm{S}_{\mathrm{train}}\gets\;\{\bm\mu_1,\ldots,\bm\mu_{N_\mathrm{train}}\}$}\Comment{initialize training set}
    \State{$\bm{S}_{\mathrm{test}}\gets\;\{\bm\mu_1,\ldots,\bm\mu_{N_\mathrm{test}}\}$}\Comment{initialize testing set}
    \State{$\bm{M}\gets\{\bm{T}_{\bm\mu_1}(\bm{g}^1),\ldots,\bm{T}_{\bm\mu_1}(\bm{g}^{n_t}),\ldots,\bm{T}_{\bm\mu_{N_{\mathrm{test}}}}(\bm{g}^1),\ldots,\bm{T}_{\bm\mu_{N_{\mathrm{test}}}}(\bm{g}^{n_t})\}$}\Comment{initialize $N_{\mathrm{test}}\cdot\,n_t$ test vectors}
    \State{$\varepsilon_{\mathrm{testfail}}\gets\varepsilon_{\mathrm{algofail}}/N_{\bm{T}}$}
    \State{$c_{\mathrm{est}}\gets{\left[\sqrt{2\lambda^{M_S}_{\min}}\mathrm{erf}^{-1}\left(\sqrt[n_t]{\varepsilon_{\mathrm{testfail}}}\right)\right]}^{-1}$}\Comment{determine error estimator factor}
    \While{$\left(\max_{t\in \bm{M}} \norm{t}_R\right) \vdot c_{\mathrm{est}} > \texttt{tol}$}
        \State{$\bm\mu_j\gets\mathrm{random.choice}(\bm{S}_{\mathrm{train}})$}
        \State{$\bm{B}\gets \bm{B}\cup\left(\bm{T}_{\bm\mu_j}(\bm{g})\right)$}
        \State{$\bm{B}\gets \mathrm{orthonormalize}\left(\bm{B}\right)$}
        \State{$\bm{M}\gets\{t-P_{\mathrm{span}(\bm{B})}t \mid t\in \bm{M}\}$}\Comment{orthogonalize test vectors to $\mathrm{span}(\bm{B})$}
    \EndWhile{}
    \State{\textbf{return} {$X^n=\mathrm{span}\left(\bm{B}\right)$}}
    \EndFunction%
    \end{algorithmic} 
\end{algorithm}

% subsubsection Heuristic randomized range finder (end)

>>>>

The contributions of the present work are given as follows.
First, as references~\cite{Smetana2016Optimal,Taddei2018Localization} do not make use of range approximation via random sampling, a suitable training strategy to construct local approximation spaces for parameterized linear problems via random sampling is discussed.
Herein, the approach given in~\cite{Taddei2018Localization}, identified as a \textit{distributed approximate POD} (see also~\cite{Himpe2018Hierarchical}), is adopted to range approximation via random sampling.
Second, a framework for the shape optimization of lattice structures is proposed.
It combines the aforementioned algorithm to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
Furthermore, the global approximation space is constructed from the local spaces using the generalized finite element method (GFEM)~\cite{Babuska2004Generalized}.

The rest of the article is organized as follows.
In~\cref{sec:method}, the building blocks of the shape optimization framework are described.
In particular, the range approximation of a parametric transfer operator is discussed.
\Cref{sec:numerical experiments} comprises the numerical experiments.
Based on the example of a graded concrete slab, the quality of the local spaces generated by method described in~\cref{sec:Randomized range finder and POD} is analyzed, the proposed ROM is validated, and results for the solution of a shape optimization problem are presented.
Finally, conclusions are given in~\cref{sec:conclusions}.

\section{METHOD}%
\label{sec:method}
The method proposed in this article is a localized pMOR framework for the shape optimization of lattice structures.
First, the auxiliary problem to model geometrical parameterizations is introduced in~\cref{sub:Auxiliary Problem}.
Second, the construction of local approximation spaces is described in~\cref{sub:Construction of local approximation spaces}.
Herein, the parametric oversampling problem and training strategy to approximate the range of the corresponding parametric transfer operator are discussed.
Third, it is outlined how a global approximation is obtained from local approximation spaces via the GFEM.
Finally, the necessity and use of hyper-reduction (in the form of empirical interpolation) to ensure an online-efficient ROM is detailed in~\cref{sub:Online efficiency}.

\subsection{Auxiliary Problem} % (fold)
\label{sub:Auxiliary Problem}
The approach adopted in this work to facilitate geometrical parameterizations belongs to the class of surface-based deformations~\cite{Botsch2010Polygon} and is based on transformations $\Phi_{\bm\mu}$ that map each material point $\bm{x}_{\mathrm{p}}$ of a parameter-independent reference or parent domain $\varOmega^{\mathrm{p}}$ to a point $\bm{x}_{\bm\mu}$ in the parameter-dependent current or physical domain $\varOmega^{\bm\mu}$.
In the context of shape optimization of lattice structures, our objective is to determine the mapping $\Phi_{\bm\mu}$ for a single unit cell, see~\cref{fig:transformationmap}, and use it to describe the change in geometry of each unit cell throughout the structure.
An auxiliary problem based on the equations of linear elasticity is solved to obtain such domain transformations, following the approach outlined in~\cite{Guo2022Learning}, and briefly repeated here for completeness.

@@

Given the transformation displacement $\bm{d}(\bm{x}_{\mathrm{p}};\bm\mu)$, the variational formulation is formulated over the parameter-independent parent domain $\varOmega^{\mathrm{p}}$ instead of the parameter-dependent physical domain $\varOmega^{\bm\mu}$.
This introduces the parameter dependence in the variational formulation as shown in the following subsections, but has the advantage that numerical integration can be carried out over the fixed parent domain and thus no re-meshing is required.
However, for each new parameter value $\bm\mu$ for which the model is to be evaluated, we first need to solve the auxiliary problem.
Nevertheless, this does not pose a problem, since the solution of the auxiliary problem is well amenable to acceleration via established MOR techniques and it is referred to~\cite{Guo2022Learning} for further details.

@@

\subsection{Construction of local approximation spaces} % (fold)
\label{sub:Construction of local approximation spaces}
In this paper, local approximation spaces are constructed by solving an oversampling problem many times for different parameter values $\bm\mu$ and different (random) boundary conditions.
It is therefore useful to cast this oversampling problem in the form of a parameter-dependent transfer operator $\bm{T}_{\bm\mu}$ that
maps the boundary function $\bm{g}$ to the solution $\bm{u}(\bm\mu)$ in the target subdomain $\varOmega_{\mathrm{in}}$.
The approximation of the range of this transfer operator is then the sought after local approximation space.

\added[id=pd]{TODO: need to introduce pull back here already because this is used in the training as well}
% First, use physical domain (suffix mu everywhere)
% Then introduce pull back and write omega_p everywhere ?

First, the global domain $\varOmega_{\mathrm{gl}}^{\bm\mu}$ and a non-overlapping domain decomposition
\begin{equation}
	\varOmega_{\mathrm{gl}}^{\bm\mu}=\cup_{i=1}^{N_{\mathrm{cells}}} \varOmega_i^{\bm\mu}
\end{equation}
is introduced.
In the context of lattice structures, each subdomain $\varOmega_i^{\bm\mu}$ corresponds to a unit cell.
Next, a coarse grid partition of the global domain and a fine grid partition of the unit cell is introduced as shown in~\added[id=pd]{TODO: add figure showing coarse grid and fine grid for a single unit cell.}
\added[id=pd]{Fig. XX} also shows one exemplary oversampling domain $\varOmega^{\bm\mu}$.
The oversampling problem then comprises the solution of the following boundary value problem.
\begin{align}
	\begin{split}
	\label{eq:osp_p}
	u &= 0\,,\\
	v &= 1
	\end{split}
\end{align}

\added[id=pd]{TODO: variational formulation oversampling problem. Maybe skip BVP in order to reduce number of eqs.}
% - setting, global domain = union of non overlapping subdomains = unit cells
% - introduction of coarse grid where each cell corresponds to unit cell subdomain
% - goal of constructing local spaces for only parts of the domain
% - definition of oversampling problem and transfer operator
% - the oversampling problem should be formulated for union of parent unit cells
% - weak formulation is not required here? Should be first introduced in section 2.3?
% - comment on number of archetypes/configurations

\added[id=pd]{TODO: add footnote about extension of Auxiliary Problem from unit cell to global domain / oversampling domain.}

% oversampling problem pulled back to parent domain
\begin{align}
	\begin{split}
		\label{eq:osp_mu}
		u &= 0 \,,\\
		v &= 1
	\end{split}
\end{align}

@@
